---
title: "Marketing Attribution Dashboard"
subtitle: "MSDM Culminating Experience Project · Trauma Resource Institute"
author: "Shayah LaCour"
date: today
format:
  dashboard:
    theme:
      - flatly
      - custom.scss
    logo: ""
    nav-buttons:
      - icon: github
        href: https://github.com/slacour2319
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(ggplot2)
library(dplyr)

# Colour palette consistent with site
clr_navy  <- "#1B4F72"
clr_terra <- "#E8825A"
clr_green <- "#2E8B6E"
clr_muted <- "#8FAABC"

channels <- c("Affiliate", "Mobile", "Paid Search", "Social Media")
colors4  <- c("#1B4F72", "#E8825A", "#2E8B6E", "#8172B3")
```

# Overview

## Row {height=25%}

```{r}
#| content: valuebox
#| title: "Total Users"
list(icon = "people-fill", color = "primary", value = 105)
```

```{r}
#| content: valuebox
#| title: "Conversions"
list(icon = "check-circle-fill", color = "success", value = 83)
```

```{r}
#| content: valuebox
#| title: "Conversion Rate"
list(icon = "graph-up-arrow", color = "warning", value = "79.0%")
```

```{r}
#| content: valuebox
#| title: "Unique Channels"
list(icon = "diagram-3-fill", color = "info", value = 4)
```

## Row {height=75%}

### Column {width=50%}

```{r}
#| title: "First Touch vs Last Touch Attribution"
comparison_df <- data.frame(
  method  = rep(c("First Touch", "Last Touch", "Markov Chain"), each = 4),
  channel = rep(channels, times = 3),
  value   = c(
    0.254, 0.418, 0.209, 0.119,
    0.388, 0.224, 0.269, 0.119,
    0.260, 0.245, 0.238, 0.257
  )
)

ggplot(comparison_df, aes(x = method, y = value, fill = channel)) +
  geom_col(position = "fill", width = 0.65) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_manual(values = colors4) +
  labs(x = NULL, y = "Share of attribution credit", fill = "Channel") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    plot.background = element_rect(fill = "white", color = NA)
  )
```

### Column {width=50%}

```{r}
#| title: "Markov Chain: Removal Effect (Drop in Conversion Probability)"
removal_df <- data.frame(
  channel = channels,
  drop    = c(0.6483, 0.6830, 0.6747, 0.6154)
)

ggplot(removal_df, aes(x = reorder(channel, drop), y = drop, fill = channel)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = scales::percent(drop, accuracy = 0.1)),
            hjust = -0.15, size = 4, fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = colors4) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 0.82)) +
  labs(x = NULL, y = "Drop in conversion probability") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    plot.background = element_rect(fill = "white", color = NA)
  )
```

# Attribution Detail

## Row

### Column {width=60%}

```{r}
#| title: "Attribution Credits by Method"
credits_df <- data.frame(
  channel = rep(channels, times = 4),
  method  = rep(c("First Touch", "Last Touch", "Markov", "Shapley"), each = 4),
  credits = c(
    34, 21, 16, 12,      # first touch
    24, 31, 22, 6,       # last touch
    21.4, 21.6, 20.5, 19.5,  # markov
    20.7, 20.6, 21.0, 20.7   # shapley
  )
)

ggplot(credits_df, aes(x = channel, y = credits, fill = method)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = c(clr_navy, clr_terra, clr_green, "#8172B3")) +
  labs(x = NULL, y = "Conversion credits", fill = "Model") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    plot.background = element_rect(fill = "white", color = NA)
  )
```

### Column {width=40%}

```{r}
#| title: "Shapley Values (Average Marginal Contribution)"
shapley_df <- data.frame(
  channel = channels,
  value   = c(0.2001, 0.1962, 0.1969, 0.1973)
)

ggplot(shapley_df, aes(x = reorder(channel, value), y = value, fill = channel)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = round(value, 4)), hjust = -0.15, size = 4) +
  coord_flip() +
  scale_fill_manual(values = colors4) +
  scale_y_continuous(limits = c(0, 0.24)) +
  labs(x = NULL, y = "Shapley value") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    plot.background = element_rect(fill = "white", color = NA)
  )
```

# Path Analysis

## Row

```{r}
#| title: "Average Conversion Rate by Path Length Quartile"
set.seed(42)
path_df <- data.frame(
  quartile = rep(c("Q1 (6–9)", "Q2 (10–17)", "Q3 (18–31)", "Q4 (32–45)"), each = 26),
  conversion = c(
    rbinom(26, 1, 0.69),
    rbinom(26, 1, 0.77),
    rbinom(26, 1, 0.84),
    rbinom(26, 1, 0.86)
  )
) |>
  group_by(quartile) |>
  summarise(rate = mean(conversion), .groups = "drop")

ggplot(path_df, aes(x = quartile, y = rate, group = 1)) +
  geom_line(color = clr_terra, linewidth = 1.5) +
  geom_point(size = 5, color = clr_navy) +
  geom_text(aes(label = scales::percent(rate, accuracy = 1)),
            vjust = -1, size = 4, fontface = "bold", color = clr_navy) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0.5, 1)) +
  labs(x = "Path length quartile (touches)", y = "Conversion rate") +
  theme_minimal(base_size = 13) +
  theme(plot.background = element_rect(fill = "white", color = NA))
```

```{r}
#| title: "Channel Touch Frequency (All Users)"
freq_df <- data.frame(
  channel = channels,
  pct     = c(28.5, 30.1, 26.4, 15.0)
)

ggplot(freq_df, aes(x = reorder(channel, pct), y = pct, fill = channel)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = paste0(pct, "%")), hjust = -0.15, size = 4.5, fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = colors4) +
  scale_y_continuous(limits = c(0, 38)) +
  labs(x = NULL, y = "% of all touchpoints") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    plot.background = element_rect(fill = "white", color = NA)
  )
```

# Key Findings

## Row

::: {.card}
### Summary of Findings

This dashboard presents marketing attribution analysis for the **Trauma Resource Institute (TRI)** as part of the MSDM Culminating Experience Project (AO #2). The goal is to measure which digital marketing channels contribute most to conversion across TRI's homepage and four program pages.

**Key conclusions:**

1. **All four channels matter.** Markov and Shapley models both show roughly equal contribution (~20–22 credits each), unlike first-touch which over-credits paid search (34) and last-touch which over-credits mobile (31).

2. **Mobile and Paid Search are structurally essential.** The removal effect analysis shows that removing either channel causes the largest drop in conversion probability (~68%), making them the highest-risk channels to cut.

3. **Social Media is undervalued by heuristics.** First-touch and last-touch both give social media the lowest credit (12 and 6 respectively), but the Markov model lifts it to near-parity (19.5 credits), suggesting it plays an important mid-funnel role that simple rules miss.

4. **Longer journeys convert better.** Users with more touchpoints have higher conversion rates, suggesting retargeting and multi-channel nurturing improve outcomes.

**Recommendation:** Use the Markov model as the primary attribution framework and Shapley values as a fairness check before making budget allocation decisions.
:::
