---
title: "Building Interactive Apps with Shiny"
subtitle: "A practical introduction for analysts who already know R"
author: "Shayah LaCour"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    code-fold: false
execute:
  echo: true
  eval: false
  warning: false
---

```{=html}
<div class="essay-header">
  <h1>⚡ Building Interactive Apps with Shiny</h1>
  <p>A practical lecture note for analysts ready to move beyond static reports</p>
</div>
```

## What is Shiny?

If you've ever finished an R analysis and thought *"I wish my manager could just drag a slider and see this update"* — that's exactly what **Shiny** solves.

Shiny is an R package developed by Posit (formerly RStudio) that lets you turn ordinary R code into a fully interactive web application — without writing a single line of JavaScript, HTML, or CSS (unless you want to). Your colleague can open a URL in a browser, adjust inputs, and see outputs update in real time, all powered by R running on a server behind the scenes.

::: {.callout-note}
## The core idea in one sentence
Shiny wraps R in a reactive framework: when an **input** changes (a slider, a dropdown, a text box), any **output** that depends on it automatically re-runs and updates.
:::

---

## The Two-Part Structure

Every Shiny app has exactly two logical pieces:

| Part | What it does | Where it lives |
|---|---|---|
| **UI** (`ui`) | Defines the layout and controls the user sees | `ui.R` or the `ui` object |
| **Server** (`server`) | Contains the R code that produces outputs | `server.R` or the `server` function |

You can write these in two separate files (`ui.R` + `server.R`) or in a single `app.R` file. For learning purposes, a single file is easiest.

---

## Your First Shiny App (10 Lines)

Here is the smallest working Shiny app. It takes a number from a slider and shows its square root.

```{r}
library(shiny)

ui <- fluidPage(
  sliderInput("n", "Pick a number", min = 1, max = 100, value = 25),
  textOutput("result")
)

server <- function(input, output) {
  output$result <- renderText({
    paste("Square root of", input$n, "is", round(sqrt(input$n), 3))
  })
}

shinyApp(ui, server)
```

::: {.callout-tip}
## Try it yourself
Paste this into RStudio and click **Run App**. You'll see a slider and a line of text. Drag the slider — notice the text updates instantly. That instant update is Shiny's reactive system doing its job.
:::

---

## Understanding Reactivity

Reactivity is what makes Shiny feel alive. The rule is simple:

> **Any output that reads `input$x` will automatically re-run whenever `input$x` changes.**

You don't call a refresh function. You don't write event listeners. Shiny tracks the dependency for you.

### The reactive graph

```
[sliderInput: n]  ──────►  [renderText: result]
```

If you had two outputs both reading `input$n`, they would both re-run when the slider moves. If you had an output that reads `input$a` and `input$b`, it re-runs when *either* changes.

### `reactive()` — sharing computed values

If multiple outputs need the same intermediate computation, don't repeat the code. Wrap it in `reactive()`:

```{r}
server <- function(input, output) {
  
  # Computed once, shared by multiple outputs
  filtered_data <- reactive({
    mtcars[mtcars$cyl == input$cyl, ]
  })
  
  output$plot  <- renderPlot({ plot(filtered_data()$wt, filtered_data()$mpg) })
  output$table <- renderTable({ filtered_data() })
}
```

Note: you call a reactive expression like a function — `filtered_data()` — with parentheses.

---

## Key UI Components

Shiny has built-in input and output widgets for almost everything you need.

### Inputs

```{r}
# Numeric input
sliderInput("year", "Select year:", min = 2018, max = 2024, value = 2022)

# Dropdown
selectInput("channel", "Channel:", choices = c("Mobile", "Paid Search", "Affiliate", "Social Media"))

# Checkbox group
checkboxGroupInput("methods", "Attribution methods:",
                   choices = c("First Touch", "Last Touch", "Markov"),
                   selected = "Markov")

# Date range
dateRangeInput("dates", "Date range:",
               start = "2024-01-01", end = Sys.Date())
```

### Outputs

```{r}
# In the UI: declare where each output goes
plotOutput("my_chart")
tableOutput("my_table")
textOutput("my_text")
verbatimTextOutput("my_code_output")

# In the server: produce each output
output$my_chart <- renderPlot({ ... })
output$my_table <- renderTable({ ... })
output$my_text  <- renderText({ ... })
```

---

## A More Realistic Example: Attribution Explorer

Here is a Shiny app closer to something you might actually use — an interactive attribution comparison tool:

```{r}
library(shiny)
library(ggplot2)
library(dplyr)

# Data
attr_data <- data.frame(
  channel = rep(c("Mobile", "Paid Search", "Affiliate", "Social Media"), 3),
  method  = rep(c("First Touch", "Last Touch", "Markov"), each = 4),
  credits = c(21, 34, 17, 10,   # first touch
              26, 15, 18, 8,    # last touch
              21, 20, 20, 21)   # markov
)

ui <- fluidPage(
  titlePanel("Attribution Model Explorer"),
  
  sidebarLayout(
    sidebarPanel(
      checkboxGroupInput(
        "methods",
        "Select attribution models:",
        choices  = unique(attr_data$method),
        selected = unique(attr_data$method)
      ),
      hr(),
      p("Adjust which models to display and compare how each model assigns
        credit to the four marketing channels.")
    ),
    
    mainPanel(
      plotOutput("attr_chart", height = "400px"),
      hr(),
      tableOutput("attr_table")
    )
  )
)

server <- function(input, output) {
  
  filtered <- reactive({
    attr_data |> filter(method %in% input$methods)
  })
  
  output$attr_chart <- renderPlot({
    ggplot(filtered(), aes(x = channel, y = credits, fill = method)) +
      geom_col(position = "dodge", width = 0.65) +
      scale_fill_brewer(palette = "Set2") +
      labs(title = "Conversion Credits by Channel and Model",
           x = NULL, y = "Credits", fill = "Model") +
      theme_minimal(base_size = 14)
  })
  
  output$attr_table <- renderTable({
    filtered() |>
      tidyr::pivot_wider(names_from = method, values_from = credits)
  })
}

shinyApp(ui, server)
```

---

## Layout Options

Shiny gives you several layout options to structure more complex apps.

### `sidebarLayout()` — most common
Controls on the left, output on the right. Used in the example above.

### `navbarPage()` — multi-tab apps
```{r}
ui <- navbarPage(
  "My App",
  tabPanel("Overview",   plotOutput("overview_plot")),
  tabPanel("Data Table", tableOutput("data_table")),
  tabPanel("About",      p("This app was built with Shiny."))
)
```

### `fluidRow()` + `column()` — custom grids
```{r}
ui <- fluidPage(
  fluidRow(
    column(4, wellPanel(sliderInput(...))),
    column(8, plotOutput("chart"))
  )
)
```

---

## How to Deploy a Shiny App

A Shiny app requires a **live R process** to run. That means you can't just put the files on a web server like a static HTML page — you need somewhere R is actually executing.

| Option | Cost | Difficulty | Best for |
|---|---|---|---|
| **shinyapps.io** (Posit) | Free tier available | Easy | Sharing with anyone |
| **Posit Connect** | Paid (org license) | Medium | Internal enterprise use |
| **Shiny Server** (self-hosted) | Free (server cost) | Hard | Full control |
| **Docker + cloud** | Pay-as-you-go | Hard | Production scale |

For class and portfolio purposes, **shinyapps.io** is the easiest path:

```{r}
# Install rsconnect and deploy
install.packages("rsconnect")
library(rsconnect)

# Authenticate once (get token from shinyapps.io)
rsconnect::setAccountInfo(name = "YOUR_ACCOUNT",
                          token = "YOUR_TOKEN",
                          secret = "YOUR_SECRET")

# Deploy
rsconnect::deployApp(appDir = "path/to/your/app/")
```

---

## Common Mistakes (and How to Fix Them)

::: {.callout-warning}
### Mistake 1: Forgetting parentheses on reactive expressions
`filtered_data` is a function — you must call it as `filtered_data()`.
:::

::: {.callout-warning}
### Mistake 2: Putting heavy computation outside `reactive()`
Code at the top level of `server` runs *once* at startup. If you need something to re-run when inputs change, it must be inside `reactive()`, `renderPlot()`, etc.
:::

::: {.callout-warning}
### Mistake 3: Using `input$x` outside a reactive context
You can only read input values inside `reactive()`, `render*()`, or `observe()`. Reading `input$x` directly in a regular function will throw an error.
:::

---

## Summary

Shiny lets you go from an R script to a shareable, interactive web app without leaving R. The mental model is:

1. **UI** describes what the user sees and interacts with.
2. **Server** contains the R logic that produces outputs.
3. **Reactivity** connects them automatically — change an input, outputs update.

For analysts already comfortable in R and ggplot2, Shiny is a natural next step. The learning curve is mostly about understanding reactivity, not about R itself.

::: {.callout-note}
## What comes next?
In the next essay, we look at **Shinylive** — a way to run Shiny apps entirely in the browser, with no server required.
:::
